---
title: "Lab 03 - Nobel laureates"
author: "Conor Lacey"
date: "01/24/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
suppressWarnings(library(tidyverse))
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are 935 observations with 26 variables in the Nobel data set. Each row 
represents an individual 

### Exercise 2

```{r nobel_living-dataframe, message = FALSE}
nobel_living<-nobel %>% 
  filter(!is.na(country) & 
           gender != "org" &
           is.na(died_date))
```

### Exercise 3

```{r country_us-variable}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r nobel_living_science-dataframe}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r country_us-bar-graph}
nobel_living_science %>% 
  ggplot(aes(y=country_us))+
  facet_wrap(~category)+
  geom_bar()+
  ylab("")
```

It appears from the bar graphs that most nobel laureates are indeed living 
in the US.

### Exercise 4

```{r born_country_us-variable}


nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )

nobel_living_science %>% count(born_country_us)
```
It appears 105 of the nobel laureates were born in the US

### Exercise 5

```{r born_country_us-bar-graph}
nobel_living_science %>% 
  ggplot(aes(y=born_country_us))+
  facet_wrap(~category)+
  geom_bar()+
  ylab("")
```

Aside from the economics category it appears the buzzfeed article is for the most part true. Many of the nobel winners were born outside the US.

### Exercise 6

